Complex Synapse Identification

### 2020-08-05 Notes/Identification.tex
basics for identification in place

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Pib}{\boldsymbol{\Pi}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\pr}{\mathbb{P}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:

@@

The different types of plasticity (\eg potentiation and depression) are indicated by superscripts $\mu,\nu$.
The type of plasticity used between times $(t-1,t)$ is $\mu(t)$.
An entire sequence of plasticity types is denoted by $\mu[1,T]$.

@@

When we don't have any prior information about the model $\M$, we fit the model by maximising the likelihood function:
%
\begin{equation}\label{eq:like}
  \pr\cond{\w^A[0,T]}{\M} = \prod_a \pr\cond{\w^a[0,T]}{\M}.
\end{equation}
%
To do this, we define an axillary function \cite{Baum1970baumwelch,Dempster2007EM}:
%
\begin{equation}\label{eq:emq}
  Q(\M,\M') = \sum_{i^A[0,T]} \pr\cond{i^A[0,T],\w^A[0,T]}{\M} \log\!\! \brk{
      \frac{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M'} }{ \pr\cond{i^A[0,T],\w^A[0,T]}{\M} } }.
\end{equation}
%
It can be shown that, if $\M'$ is chosen to maximise this quantity (which is easier than maximising the likelihood), the it will have a higher likelihood than $\M'$.
In effect, $\M$ is used to estimate the hidden state trajectories, which are then used to re-estimate the model.
This can be done iteratively to find a local maximum of the likelihood.
If this is repeated several times with random initialisation, we can hope to find the global maximum.

When we have several sequences, we can write
%
\begin{equation}\label{eq:emqseveral}
\begin{gathered}
  Q(\M,\M') = \sum_{i^A[0,T]}  \prod_b \pr\cond{i^b[0,T],\w^b[0,T]}{\M} \sum_a  \log\!\!  \brk{
      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
   = \sum_a \prod_{b \neq a} \pr\cond{\w^b[0,T]}{\M}\! \sum_{i^a[0,T]}\!\! \pr\cond{i^a[0,T],\w^a[0,T]}{\M} \log\!\!  \brk{
      \frac{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M'} }{ \pr\cond{i^a[0,T],\w^a[0,T]}{\M} } } \\
   = \pr\cond{\w^A[0,T]}{\M} \sum_a \frac{Q^a(\M,\M')}{\pr\cond{\w^a[0,T]}{\M}}.
\end{gathered}
\end{equation}
%

@@

Suppose we have a single sequence of plasticity events.
Then the joint likelihood of a trajectory and sequence observed synaptic weights is given by
%
\begin{equation}\label{eq:trajlike}
\begin{aligned}
  \pr\cond{t[0,T], \w[0,T]}{\M} &= \pib_{i(0)} \, \Pib_{i(0)}(0) 
        \prod_{t=1}^{T} \M^{\mu(t)}_{i(t-1)i(t)} \, \Pib_{i(t)}(t),\\
  \text{where}\quad
  \Pib_i(t) &= \delta_{\w_{i}\w(t)}
\end{aligned}
\end{equation}
%
We wish to maximise \eqref{eq:emq} subject to normalisation constraints.
Therefore we maximise the Lagrangian:
%
\begin{equation}\label{eq:lagrangesingle}
  \CL = Q(\M,\M') + \lambda\prn{1-\sum_i\pib'_i} + \sum_{\mu i} \lambda^\mu_i\prn{1-\sum_j\M'^\mu_{ij}}. 
\end{equation}
%
We find
%
\begin{equation}\label{eq:singlediffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \frac{N^\mu_{ij}}{\M'^\mu_{ij}} - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \frac{\alpha_i(0) \, \beta_i(0)}{\pib'_{i}} - \lambda,
\end{aligned}
\end{equation}
%
where the numerator $N^\mu_{ij}$ and the forward/backward variables $\alpha$ and $\beta$ are defined as:
%
\begin{equation}\label{eq:alphabeta}
\begin{aligned}
  N^\nu_{jk} &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \pr\cond{i[t-1,t]=(jk),\w[0,T]}{\M} \\
     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \alpha_j(t-1) \, \beta_k(t) \, \M^{\nu}_{jk} \, \Pib_k(t) ,\\
  \alpha_j(t) &= \pr\cond{i(t)=j,\w[0,t]}{\M} \\
     &= \sum_{i[0,t]} \pib_{i(0)}(0) \, \Pib_{i(0)}(0) \prod_{s=1}^{t} \M^{\mu(s)}_{i(s-1)i(s)} \, \Pib_{i(s)}(s) \,\delta_{i(t)j}, \\
  \beta_j(t) &= \pr\cond{i(t)=j,\w[t+1,T]}{\M} \\
     &= \sum_{i[t,T]} \delta_{i(t)j} \prod_{s=t+1}^{T} \M^{\mu(s)}_{i(s-1)i(s)} \, \Pib_{i(s)}(s).
\end{aligned}
\end{equation}
%
Note that, whilst disjoint sequences $\w[s,s']$ and $\w[t,t']$ are not independent, they are conditionally independent given the state at any time between the two intervals.
Therefore:
%
\begin{equation}\label{eq:stateprob}
  \alpha_j(t) \, \beta_j(t) = \pr\cond{i(t)=j,\w[0,T]}{\M}.
\end{equation}
%

These quantities usually suffer from underflow.
This can be avoided by using normalised forward/backward variables \cite{Zhai2003HMMnorm}:
%
\begin{equation}\label{eq:normalphabeta}
\begin{aligned}
  \tilde{\alpha}_i(t) &= \alpha_i(t) \prod_{s=0}^{t} \eta(s), \\
  \tilde{\beta}_i(t) &= \beta_i(t) \prod_{s=t+1}^{T} \eta(s),
\end{aligned}
\end{equation}
%
where $\eta[0,T]$ is chosen so that all of the $\tilde{\alpha}[0,T]$ are normalised: $\sum_i \tilde{\alpha}_i(t) = 1$.
These quantities have the interpretations:
%
\begin{equation}\label{eq:normintepr}
\begin{aligned}
  \tilde{\alpha}_j(t) &= \pr\cond{i(t)=j}{\w[0,t],\M}, \\
  \eta(t) &= \pr\cond{\w(t)}{\w[0,t-1],\M}, \\
  \prod_{s=0}^t \eta(t) &= \pr\cond{\w[0,t]}{\M}.
\end{aligned}
\end{equation}
%
Note that the new backward variables $\tilde{\beta}$ are not normalised and have no particularly interesting interpretation.
However:
%
\begin{equation}\label{eq:normstateprob}
\begin{aligned}
  \widetilde{N}^\nu_{jk} &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \pr\cond{i[t-1,t]=(jk)}{\w[0,T],\M} \\
     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, \tilde{\alpha}_i(t-1) \, \tilde{\beta}_j(t) \, \eta(t) \, \M^{\nu}_{i(t-1)i(t)} \, \Pib_k(t) ,\\
  \tilde{\alpha}_j(t) \, \tilde{\beta}_j(t) &= \pr\cond{i(t)=j}{\w[0,T],\M}.
\end{aligned}
\end{equation}
%
Then we can write
%
\begin{equation}\label{eq:singlediffsnorm}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \widetilde{N}^\mu_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w[0,T]}{\M} \, \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \pib'_{i} } - \lambda,
\end{aligned}
\end{equation}
%
The factors of $\pr\cond{\w[0,T]}{\M}$ can be absorbed by a redefinition of the Lagrange multipliers.

@@

When we have multiple sequences of plasticity events, we will still want to maximise the Lagrangian \eqref{eq:lagrangesingle}, but now we will use the auxiliary function \eqref{eq:emqseveral}.
This results in
%
\begin{equation}\label{eq:multidiffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
    \frac{ N^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
    &= \pr\cond{\w^A[0,T]}{\M} \, \frac{ \sum_a \widetilde{N}^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr\cond{\w^A[0,T]}{\M} \,\sum_a \frac{1}{\pr\cond{\w^a[0,T]}{\M}} \,
    \frac{ \alpha^a_i(0) \, \beta^a_i(0) }{ \pib'_{i} } - \lambda,\\
    &= \pr\cond{\w^A[0,T]}{\M} \,
    \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \pib'_{i} } - \lambda.
\end{aligned}
\end{equation}
%
Once again, we can absorb the factors of $\pr\cond{\w[0,T]}{\M}$ by a redefinition of the Lagrange multipliers.

@@

When we have a prior distribution for the model, $\pr(\M)$, instead of maximising the likelihood \eqref{eq:like}, we can maximise the posterior:
%
\begin{equation}\label{eq:posterior}
  \pr\cond{\M}{\w[0,T]} = \frac{ \pr(\w[0,T],\M) }{ \pr(\w[0,T) } = \frac{ \pr\cond{\w[0,T]}{\M} \pr(\M) }{ \pr(\w[0,T) }.
\end{equation}
%
As the denominator is independent of the model, this is equivalent to maximising the numerator, \ie the joint distribution.

Using the same trick as before \eqref{eq:emq}, we can define an auxilliary function
%
\begin{equation}\label{eq:emqprior}
\begin{aligned}
  \widetilde{Q}(\M,\M') &= \sum_{i^A[0,T]} \pr\!\prn{i^A[0,T],\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M'} }{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M} } }\\
      &= \pr\!\prn{\M} Q(\M,\M') + \pr\!\prn{\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{\M'} }{ \pr\!\prn{\M} } }.
\end{aligned}
\end{equation}
%

@@

Here we use the off-diagonal $\CL^1$ norm as a penalty:
%
\begin{equation}\label{eq:L1penalty}
  E^\mu_{ij} = (1-\delta_{ij}) \M^\mu_{ij}.
\end{equation}
%
Then the condition for a maximum \eqref{eq:priordiffs} reads
%
\begin{equation}\label{eq:L1priormax}
\begin{aligned}
  \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - (1-\delta_{ij})\beta - \lambda^\mu_{i} &= 0, \\
   \M'^\mu_{ij} &= \frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+(1-\delta_{ij})\beta}.
\end{aligned}
\end{equation}
%
Demanding normalisation gives
%
\begin{multline}\label{eq:L1norm}
  \sum_{j \neq i}\frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+\beta} +\frac{\widetilde{N}^{A\mu}_{ii}}{\lambda^\mu_{i}} =0, \\
  \implies \quad
  \lambda^\mu_{i} = \half \brk{ \prn{\sum_j \widetilde{N}^{A\mu}_{ij} -\beta} + 
     \sqrt{\prn{\sum_j \widetilde{N}^{A\mu}_{ij} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}}}
\end{multline}
%
Which produces the update rule
%
\begin{equation}\label{eq:L1update}
\begin{aligned}
  \M'^\mu_{ij} &= \frac{ 2\widetilde{N}^{A\mu}_{ij} }{ \prn{\sum_k \widetilde{N}^{A\mu}_{ik} +\beta} +
     \sqrt{\prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}} } 
     \qquad \text{for } i\neq j, \\
  \M'^\mu_{ii} &= \frac{ 2\widetilde{N}^{A\mu}_{ii} }{ \prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta} +
     \sqrt{\prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}} }.
\end{aligned}
\end{equation}
%

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\usepackage{mathtools}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Pib}{\boldsymbol{\Pi}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\pr}{\mathbb{P}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:

@@

The different types of plasticity (\eg potentiation and depression) are indicated by superscripts $\mu,\nu$.
The type of plasticity used between times $(t,t+1)$ is $\mu(t)$.
An entire sequence of plasticity types is denoted by $\mu[0,T-1]$.

@@

When we don't have any prior information about the model $\M$, we fit the model by maximising the likelihood function:
%
\begin{equation}\label{eq:like}
\begin{aligned}
  \pr^A_{[0,T]}\cond{\w}{\mu,\M}
    &\equiv \pr\cond{\w^A[0,T]}{\mu^A[0,T-1],\M}  \\
    &= \prod_a \pr\cond{\w^a[0,T]}{\mu^a[0,T-1],\M}
      \equiv \prod_a \pr^a_{[0,T]}\cond{\w}{\mu,\M}.
\end{aligned}
\end{equation}
%
To do this, we define an axillary function \cite{Baum1970baumwelch,Dempster2007EM}:
%
\begin{equation}\label{eq:emq}
  Q(\M,\M') = \sum_{{i^A[0,T]}} 
      \pr^A_{[0,T]}\cond{i,\w}{\mu,\M} \log\!\! \brk{
      \frac{ \pr^A_{[0,T]}\cond{i,\w}{\mu,\M'} }
           { \pr^A_{[0,T]}\cond{i,\w}{\mu,\M} } }.
\end{equation}
%
It can be shown that, if $\M'$ is chosen to maximise this quantity (which is easier than maximising the likelihood), the it will have a higher likelihood than $\M'$.
In effect, $\M$ is used to estimate the hidden state trajectories, which are then used to re-estimate the model.
This can be done iteratively to find a local maximum of the likelihood.
If this is repeated several times with random initialisation, we can hope to find the global maximum.

When we have several sequences, we can write
%
\begin{equation}\label{eq:emqseveral}
\begin{aligned}
  Q(\M,\M') &= \sum_{i^A[0,T]} \prod_b 
    \pr^b_{[0,T]}\cond{i,\w}{\mu,\M} \sum_a  \log\!\!  \brk{
      \frac{ \pr^a_{[0,T]}\cond{i,\w}{\mu,\M'} }
           { \pr^a_{[0,T]}\cond{i,\w}{\mu,\M} } } \\
   &= \sum_a \prod_{b \neq a} 
   \pr^b_{[0,T]}\cond{\w}{\mu,\M}\! 
   \sum_{i^a[0,T]}\!\! \pr^a_{[0,T]}\cond{i,\w}{\mu,\M} \log\!\!  \brk{
      \frac{ \pr^a_{[0,T]}\cond{i,\w}{\mu,\M'} }
           { \pr^a_{[0,T]}\cond{i,\w}{\mu,\M} } } \\
   &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \sum_a 
      \frac{Q^a(\M,\M')}{\pr^a_{[0,T]}\cond{\w}{\mu,\M}}.
\end{aligned}
\end{equation}
%

@@

Suppose we have a single sequence of plasticity events.
Then the joint likelihood of a trajectory and sequence observed synaptic weights is given by
%
\begin{equation}\label{eq:trajlike}
\begin{aligned}
  \pr\cond{i[0,T], \w[0,T]}{\mu[0,T-1],\M} &= \pib_{i(0)} \, \Pib_{i(0)}(0) 
        \prod_{t=0}^{T-1} \M^{\mu(t)}_{i(t)i(t+1)} \, \Pib_{i(t+1)}(t+1),\\
  \text{where}\quad
  \Pib_i(t) &= \delta_{\w_{i}\w(t)}
\end{aligned}
\end{equation}
%
We wish to maximise \eqref{eq:emq} subject to normalisation constraints.
Therefore we maximise the Lagrangian:
%
\begin{equation}\label{eq:lagrangesingle}
  \CL = Q(\M,\M') + \lambda\prn{1-\sum_i\pib'_i} + \sum_{\mu i} \lambda^\mu_i\prn{1-\sum_j\M'^\mu_{ij}}. 
\end{equation}
%
We find
%
\begin{equation}\label{eq:singlediffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \frac{N^\mu_{ij}}{\M'^\mu_{ij}} - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \frac{\alpha_i(0) \, \beta_i(0)}{\pib'_{i}} - \lambda,
\end{aligned}
\end{equation}
%
where the numerator $N^\mu_{ij}$ and the forward/backward variables $\alpha$ and $\beta$ are defined as:
%
\begin{equation}\label{eq:alphabeta}
\begin{aligned}
  N^\nu_{jk} &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, 
        \pr\cond{i[t,t+1]=(j,k),\w[0,T]}{\mu[0,T-1],\M} \\
     &= \sum_{t=1}^{T} \delta_{\mu(t)\nu} \, 
        \alpha_j(t) \, \beta_k(t+1) \, \M^{\nu}_{jk} \, \Pib_k(t+1) ,\\
  \alpha_j(t) &= \pr\cond{i(t)=j,\w[0,t]}{\mu[0,t-1],\M} \\
     &= \sum_{i[0,t]} \pib_{i(0)}(0) \, \Pib_{i(0)}(0) \prod_{s=0}^{t-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib_{i(s+1)}(s+1) \,\delta_{i(t)j}, \\
  \beta_j(t) &= \pr\cond{\w[t+1,T]}{i(t)=j,\mu[t,T-1],\M} \\
     &= \sum_{i[t,T]} \delta_{i(t)j} \prod_{s=t}^{T-1} \M^{\mu(s)}_{i(s)i(s+1)} \, \Pib_{i(s+1)}(s+1).
\end{aligned}
\end{equation}
%
Note that, whilst disjoint sequences $\w[s,s']$ and $\w[t,t']$ are not independent, they are conditionally independent given the state at any time between the two intervals.
Therefore:
%
\begin{equation}\label{eq:stateprob}
  \alpha_j(t) \, \beta_j(t) = \pr\cond{i(t)=j,\w[0,T]}{\M}.
\end{equation}
%

These quantities usually suffer from underflow.
This can be avoided by using normalised forward/backward variables \cite{Zhai2003HMMnorm}:
%
\begin{equation}\label{eq:normalphabeta}
\begin{aligned}
  \tilde{\alpha}_i(t) &= \alpha_i(t) \prod_{s=0}^{t} \eta(s), \\
  \tilde{\beta}_i(t) &= \beta_i(t) \prod_{s=t+1}^{T} \eta(s),
\end{aligned}
\end{equation}
%
where $\eta[0,T]$ is chosen so that all of the $\tilde{\alpha}[0,T]$ are normalised: $\sum_i \tilde{\alpha}_i(t) = 1$.
These quantities have the interpretations:
%
\begin{equation}\label{eq:normintepr}
\begin{aligned}
  \tilde{\alpha}_j(t) &= \pr\cond{i(t)=j}{\w[0,t],\mu[0,t-1],\M}, \\
  \eta(t)\inv &= \pr\cond{\w(t)}{\w[0,t-1],\mu[0,t-1],\M}, \\
  \brk{\prod_{s=0}^t \eta(s)}\inv &= \pr\cond{\w[0,t]}{\mu[0,t-1],\M}, \\
  \brk{\prod_{s=t_0}^{t_1} \eta(s)}\inv &= 
          \pr\cond{\w[t_0,t_1]}{\w[0,t_0-1],\mu[0,t_1-1],\M}.
\end{aligned}
\end{equation}
%
Note that the new backward variables $\tilde{\beta}$ are not normalised and have no particularly interesting interpretation, beyond the following:
%
\begin{equation}\label{eq:normstateprob}
\begin{aligned}
  \tilde{\alpha}_j(t) \, \tilde{\beta}_j(t) &= \pr\cond{i(t)=j}{\w[0,T],\mu[0,T-1],\M}.
\end{aligned}
\end{equation}
%
Then we can write
%
\begin{equation}\label{eq:singlediffsnorm}
\begin{aligned}
  \widetilde{N}^\nu_{jk} &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, 
      \pr\cond{i[t,t+1]=(jk)}{\w[0,T],\mu[0,T-1],\M} \\
     &= \sum_{t=0}^{T-1} \delta_{\mu(t)\nu} \, 
      \tilde{\alpha}_j(t) \, \tilde{\beta}_k(t+1) \, \eta(t+1) \, 
      \M^{\nu}_{jk} \, \Pib_k(t+1) ,\\
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr_{[0,T]}\cond{\w}{\mu,\M} \, 
      \frac{ \widetilde{N}^\mu_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr_{[0,T]}\cond{\w}{\mu,\M} \, 
      \frac{ \tilde{\alpha}_i(0) \, \tilde{\beta}_i(0) }{ \pib'_{i} } - \lambda,
\end{aligned}
\end{equation}
%
The factors of $\pr_{[0,T]}\cond{\w}{\mu,\M}$ can be absorbed by a redefinition of the Lagrange multipliers.

@@

When we have multiple sequences of plasticity events, we will still want to maximise the Lagrangian \eqref{eq:lagrangesingle}, but now we will use the auxiliary function \eqref{eq:emqseveral}.
This results in
%
\begin{equation}\label{eq:multidiffs}
\begin{aligned}
  \pdiff{\CL}{\M'^\mu_{ij}} &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \,\sum_a \frac{1}{\pr^a_{[0,T]}\cond{\w}{\mu,\M}} \,
    \frac{ N^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
    &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \, \frac{ \sum_a \widetilde{N}^{a\mu}_{ij} }{ \M'^\mu_{ij} } - \lambda^\mu_{i}, \\
  \pdiff{\CL}{\pib'_{i}} &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \,\sum_a \frac{1}{\pr^a_{[0,T]}\cond{\w}{\mu,\M}} \,
    \frac{ \alpha^a_i(0) \, \beta^a_i(0) }{ \pib'_{i} } - \lambda,\\
    &= \pr^A_{[0,T]}\cond{\w}{\mu,\M} \,
    \frac{ \sum_a \tilde{\alpha}^a_i(0) \, \tilde{\beta}^a_i(0) }{ \pib'_{i} } - \lambda.
\end{aligned}
\end{equation}
%
Once again, we can absorb the factors of $\pr^A_{[0,T]}\cond{\w}{\mu,\M}$ by a redefinition of the Lagrange multipliers.

@@

When we have a prior distribution for the model, $\pr(\M)$, instead of maximising the likelihood \eqref{eq:like}, we can maximise the posterior:
%
\begin{equation}\label{eq:posterior}
  \pr\cond{\M}{\w[0,T]} = \frac{ \pr(\w[0,T],\M) }{ \pr(\w[0,T]) } = \frac{ \pr\cond{\w[0,T]}{\M} \pr(\M) }{ \pr(\w[0,T]) }.
\end{equation}
%
As the denominator is independent of the model, this is equivalent to maximising the numerator, \ie the joint distribution.

Using the same trick that was used for \eqref{eq:emq}, we can define an auxilliary function
%
\begin{equation}\label{eq:emqprior}
\begin{aligned}
  \widetilde{Q}(\M,\M') &= \sum_{i^A[0,T]} \pr\!\prn{i^A[0,T],\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M'} }{ \pr\!\prn{i^A[0,T],\w^A[0,T],\M} } }\\
      &= \pr\!\prn{\M} Q(\M,\M') + \pr\!\prn{\w^A[0,T],\M} \log\!\! \brk{
      \frac{ \pr\!\prn{\M'} }{ \pr\!\prn{\M} } }.
\end{aligned}
\end{equation}
%

@@

Here we use the off-diagonal $\CL^1$ norm as a penalty:
%
\begin{equation}\label{eq:L1penalty}
  E^\mu_{ij} = (1-\delta_{ij}) \M^\mu_{ij}.
\end{equation}
%
Then the condition for a maximum \eqref{eq:priordiffs} reads
%
\begin{equation}\label{eq:L1priormax}
\begin{aligned}
  \frac{ \widetilde{N}^{A\mu}_{ij} }{ \M'^\mu_{ij} }  - (1-\delta_{ij})\beta - \lambda^\mu_{i} &= 0, \\
   \M'^\mu_{ij} &= \frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+(1-\delta_{ij})\beta}.
\end{aligned}
\end{equation}
%
Demanding normalisation gives
%
\begin{multline}\label{eq:L1norm}
  \sum_{j \neq i}\frac{\widetilde{N}^{A\mu}_{ij}}{\lambda^\mu_{i}+\beta} +\frac{\widetilde{N}^{A\mu}_{ii}}{\lambda^\mu_{i}} = 1, \\
  \implies \quad
  \lambda^\mu_{i} = \half \brk{ \prn{\sum_j \widetilde{N}^{A\mu}_{ij} -\beta} + 
     \sqrt{\prn{\sum_j \widetilde{N}^{A\mu}_{ij} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}}}
\end{multline}
%
Which produces the update rule
%
\begin{equation}\label{eq:L1update}
\begin{aligned}
  \M'^\mu_{ij} &= \frac{ 2\widetilde{N}^{A\mu}_{ij} }{ \prn{\sum_k \widetilde{N}^{A\mu}_{ik} +\beta} +
     \sqrt{\prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}} } 
     \qquad \text{for } i\neq j, \\
  \M'^\mu_{ii} &= \frac{ 2\widetilde{N}^{A\mu}_{ii} }{ \prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta} +
     \sqrt{\prn{\sum_k \widetilde{N}^{A\mu}_{ik} -\beta}^2 +4\beta\widetilde{N}^{A\mu}_{ii}} }.
\end{aligned}
\end{equation}
%
